---
title: "Hypothesis testing errors: opportunity cost"
output: html_notebook
---

```{r}
# Load packages
library(ggplot2)
library(tidyverse)
library(infer)
```

```{r}
opportunity <- structure(list(decision = structure(c(1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L
), .Label = c("buyDVD", "nobuyDVD"), class = "factor"), group = structure(c(1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 2L), .Label = c("control", "treatment"), class = "factor")), class = "data.frame", row.names = c(NA, 
-150L))

opportunity
```



# Example: opportunity cost

## Summarizing opportunity cost (1)
```{r}
# Tabulate the data
opportunity %>%
  count(decision, group)
```


```{r}
# Find the proportion who bought the DVD in each group
opportunity %>%
  group_by(group) %>%
  summarize(buy_prop = mean(decision == "buyDVD"))
```
Okay, so about 75% of the control group bought the DVD and about 55% of the treatment group (i.e. The group that was reminded that the money could be saved) bought the DVD. Interesting!


## Plotting opportunity cost (1)

Again, interest is in whether the treatment and control groups were equally likely to buy a DVD after reading the experimental statements. Here, you'll create a barplot to visualize the difference in proportions between the treatment and control groups.

```{r}
# Plot group, filled by decision
ggplot(opportunity, aes(x = group, fill = decision)) + 
  # Add a bar layer, with position "fill"
  geom_bar(position = "fill") +
  ggtitle("oportunity cost")
```


## Randomizing opportunity cost
As in Chapter 2 Exercise 5, you will permute the data to generate a distribution of differences as if the null hypothesis were true.

In the study, the number of individuals in each of the control and treatment groups is fixed. Additionally, when you assume that the null hypothesis is true—that is, the experiment had no effect on the outcome of buying a DVD—it is reasonable to infer that the number of individuals who would buy a DVD is also fixed. That is, 97 people were going to buy a DVD regardless of which treatment group they were in.

Using the new data and the methods from the previous chapter, create a randomization distribution of the difference in proportions calculated on permuted data.

```{r}
# Calculate the observed difference in purchase rate
diff_orig <- opportunity %>%
  # Group by group
  group_by(group) %>%
  # Calculate proportion deciding to buy a DVD
  summarize(prop_buy = mean(decision == "buyDVD")) %>%
  # Calculate difference between groups
  summarize(stat = diff(prop_buy)) %>% 
  pull()

diff_orig
```


```{r}
# Create data frame of permuted differences in purchase rates
opp_perm <- opportunity %>%
  # Specify decision vs. group, where success is buying a DVD
  specify(decision ~ group, success = "buyDVD") %>%
  # Set the null hypothesis to independence
  hypothesize(null = "independence") %>%
  # Generate 1000 reps of type permute
  generate(reps = 1000, type = "permute") %>%
  # Calculate the summary stat difference in proportions
  calculate(stat = "diff in props", order = c("treatment", "control"))
    
# Review the result
opp_perm
```

```{r}
# Using the permuation data, plot stat
ggplot(opp_perm, aes(x = stat)) + 
  geom_histogram(binwidth = 0.005) +
  geom_vline(aes(xintercept = diff_orig), color = "red") + 
  ggtitle("Randomized distribution of difference in proportions")
```


## Summarizing opportunity cost (2)

Now that you've created the randomization distribution, you'll use it to assess whether the observed difference in proportions is consistent with the null difference. You will measure this consistency (or lack thereof) with a p-value, or the proportion of permuted differences less than or equal to the observed difference.

```{r}
# Visualize the statistic 
opp_perm %>%
  visualize(obs_stat = diff_orig, direction = "less")

# Calculate the p-value using `get_p_value`
opp_perm %>%
  get_p_value(obs_stat = diff_orig, direction = "less")

# Calculate the p-value using `summarize`
opp_perm %>%
  summarize(p_value = mean(stat <= diff_orig))
```


p-value is so small (0.006) that is not int he range of the above plot.

## Opportunity cost conclusion

 We can confidently say the different messaging caused the students to change their buying habits, since they were randomly assigned to treatment and control groups. Let's continue.

# Errors and their consecuences






# Summary of opportunity costs